Reproducible Research - Peer Assessment: # 1
-----------------------------------

Assignment work by: csbyjw
       on: 2014-05-18

## Read in the data (activity.csv)

This code sets the working directory for this assignment repository then attempts
to load the file activity.csv from the top level of the directory

```{r}
library(ggplot2)
setwd("~/Documents/Coursera/Data Science/code")
activity <- read.csv("activity.csv", header=TRUE, stringsAsFactors=FALSE)
```

## Reshaping the data

Agg the data by day and by interval for later computations
then store the result in .bydate and .byinterval frames. Calc the mean, median, 
and total for each agg then store the result in .mn, .md, and .total variables

```{r}
activity.byday <- aggregate(steps~date, data=activity, FUN=sum)
activity.byinterval <- aggregate(steps~interval, data=activity, FUN=mean)
str(activity.byday)
str(activity.byinterval)
```

## What is the mean total number of steps per day?

Histogram of frequency of steps:
```{r}
qplot(steps, data=activity.byday, main="Histogram of Steps per Day", xlab="Total Steps")
```

Calc the mean and the median:
```{r}
mean_steps <- mean(activity.byday$steps, na.rm=TRUE)
median_steps <- median(activity.byday$steps, na.rm=TRUE)
```

The mean number of steps per day is `r format(mean_steps, scientific=FALSE)` and the median is `r format(median_steps, scientific=FALSE)`.

## What is the avg daily activity pattern?

Line plot of avg steps per interval:
```{r}
qplot(interval,steps,data=activity.byinterval)+geom_line()
```

Find the interval with the highest avg steps:
```{r}
max_interval <- activity.byinterval[activity.byinterval$steps==max(activity.byinterval$steps),'interval']
```

The interval with the most steps is `r max_interval`

## Imputing missing values

Calculate number of NA rows:
```{r}
missing <- sum(is.na(activity$steps))
```

There are `r missing` rows with missing data.
Data related to: 11/30 is missing, causing a problem when plotting by date.
Replacing missing values with avg steps for this date
```{r}
activity.filled <- activity
for (i in 1:nrow(activity.filled)) {
    r <- activity.filled[i,]
    if (is.na(r$steps)) {
        activity.filled[i,"steps"] <- 
            activity.byinterval[activity.byinterval$interval==r$interval, "steps"]
    }
}
```

Agg filled data by day and plot the histogram.
```{r}
activity.filled.byday <- aggregate(steps~date, data=activity.filled, FUN=sum)
qplot(steps,data=activity.filled.byday,main="Histogram of Steps per Day",xlab="Total Steps")
```

Calc mean and median:
```{r}
mean_steps_f <- mean(activity.filled.byday$steps, na.rm=TRUE)
median_steps_f <- median(activity.filled.byday$steps, na.rm=TRUE)
```

The mean number of steps per day is `r format(mean_steps_f, scientific=FALSE)` and
the median is `r format(median_steps_f, scientific=FALSE)`. 

### Result: The values are skewed due to the imputation

## Are the activity patterns different between weekends and weekdays?

Create a new factor for weekend vs. weekday on activity.filled
```{r}
activity.filled$weekend <- factor(weekdays(as.Date(activity.filled$date)) %in% c('Saturday','Sunday'),labels=c('Weekday','Weekend'))
```

Plot the results of mean by interval, faceted by weekend factor. 
```{r}
activity.filled.byintv_wday <- aggregate(steps~interval+weekend, data=activity.filled, FUN=mean)
qplot(interval,steps,data=activity.filled.byintv_wday,facets=weekend~.,main="Average Steps per Interval",xlab="Total Steps")+geom_line()
```
